\section{Problems}

\begin{prob}
    Let $X$ be a completely regular space.
    Show that $X$ is connected if and only if its Stone-\v{C}ech compactification is connected.
\end{prob}
\begin{sol}
    $\beta X$ is connected when $X$ is connected, since the closure of $X$ in $\beta X$ is $\beta X$ and the connectedness is preserved under closing subspaces.
    Thus, it remains to show that $X$ is connected when $\beta X$ is connected.
    Suppose $X$ is disconnected and write $X=A\sqcup B$, where $A$ and $B$ are nonempty disjoint open subsets of $X$.
    Let $f: X\rightarrow \{0, 1\}$ be the function defined $f(a)=0$ for all $a\in A$ and $f(b)=1$ for all $b\in B$, and let $\beta f: \beta X\rightarrow \{0, 1\}$ be the continuous function induced by $f$. (Note that $\{0, 1\}$ is a finite set, so it is a compact Hausdorff space.)
    Since $\beta X$ is connected, the image of $\beta f$ is $[0, 1]$ in $\bb{R}$, a contradiction.
    Therefore, $X$ is connected if $\beta X$ is connected.
\end{sol}

\begin{prob}
    Let $X$ be a discrete space and let $(\beta X, \iota)$ be the Stone-\v{C}ech compactification of $X$.
    \begin{enumerate}
        \item[(a)]
        {
            Show that if $A$ is a subset of $X$ then the closures of $A$ and $X\setminus A$ in $\beta X$ are disjoint.
        }
        \item[(b)]
        {
            Show that if $U$ is an open subset of $\beta X$ then so is its closure in $\beta X$.
        }
        \item[(c)]
        {
            Show that $\beta X$ is totally disconnected.
        }
    \end{enumerate}
\end{prob}
\begin{sol}
    
\end{sol}

\begin{prob}
    Let $X$ be a completely regular space.
    \begin{enumerate}
        \item[(a)]
        {
            Suppose $X$ is normal and identify $X$ as a subset of $\beta X$ (or choose $\beta X$ as a compactification of $X$ satisfying the universal property of the Stone-\v{C}ech compactification).
            Show that $y$ is not a limit point of $X$ in $\beta X$.
        }
        \item[(b)]
        {
            Show that $\beta X$ is not metrizable if $X$ is completely regular but not compact.
        }
    \end{enumerate}
\end{prob}
\begin{sol}
    
\end{sol}