\section{The Stone-\v{C}ech compactification}

Throughout this section, we assume that $X$ is a completely regular space, unless stated otherwise.

Finding a compactification $Y$ of a completely regular space $X$ to which every continuous map from $X$ into $\bb{R}$ is continuously extended is a basic problem.
Regarding this, if a given real-valued function on $X$ were to be extended, then the function must have been bounded.

The idea to find such a compactification of $X$ is to apply \cref{emb&cptf}.
To be precise, we first find an appropriate embedding of $X$ in terms of all the bounded continuous functions on $X$.
\cref{emb&cptf} then asserts the existence of an extended embedding of a compactification of $X$, which surely is in terms of the ``extended'' bounded continuous functions on $Y$.
This idea is applicable, since we know from the embedding theorem that a completely regular space can be embedded into $[0, 1]^I$ for some $I$, which is a compact space.

Write $I=C_b^0(X, \bb{R})$ for convinience.
For each $\alpha\in I$, let $B_\alpha=[\inf(\alpha), \sup(\alpha)]\subset\bb{R}$, and define
\begin{align*}
    B=\prod_{\alpha\in I}B_\alpha.
\end{align*}
Since $I$ separates points of $X$ from closed subsets of $X$, the map $F: X\rightarrow B$ defined by $F=(f)_{f\in I}$ is an embedding of $X$ into the compact Hausdorff space $B$.
By \cref{emb&cptf}, there is a unique compactification $Y$ of $X$ such that $F$ extends to an embedding $F_*$ of $Y$ into $B$; namely, the compactification of $X$ induced by the embedding $F$.
The $f$-component $f_*$ of $F_*$ is a desired extension of $f$, since $f_*=\pi_f\circ F_*$ is continuous.
This proves the existence of a continuous extension of $f$ to a compactification of $X$.
Remark that the uniqueness of a continuous extension of $f$ to a compactification of $X$ is guaranteed by the assumption that the codomain is a Hausdorff space.
\begin{equation*}
\begin{tikzcd}[row sep=large, column sep=huge]
    &
    X
    \arrow[dl, "\textsf{inclusion}"{sloped}, hook]
    \arrow[d, "F"]
    \arrow[dr, "f"]
    &
    \\
    Y
    \arrow[r, "F_*"', hook]
    &
    B
    \arrow[r, "\pi_f"']
    &
    \bb{R}
\end{tikzcd}
\end{equation*}

To sum up, by considering an embedding $F$ with regard to $C_b^0(X, \bb{R})$ and considering the compactification $Y$ of $X$ induced by $F$, we could show that every function in $C_b^0(X, \bb{R})$ extends uniquely to a continuous function on $Y$.
The above observation is summarized as the following proposition:
\begin{prop}\label{SC-cptf_ver1}
    Let $X$ be a completely regular space.
    Then there is a compactification $Y$ of $X$, to which every function $f\in C_0^b(X, \bb{R})$ extends continuously to $Y$ uniquely.
\end{prop}

We wish to consider a compactification of a completely regular space which satisfies a more general property.
Such compactification would be defined in terms of a universal property.

\begin{defi}[The Stone-\v{C}ech compactification of a topological space]\label{SC-cptf}
    Let $X$ be a topological space (not necessarily a completely regular space).
    A pair $(\beta(X), \iota)$ of a compact Hausdorff space $\beta(X)$ and a continuous map $\iota: X\rightarrow \beta(X)$ is called a Stone-\v{C}ech compactification of $X$, if the pair satisfies the following universal property:
    For any compact Hausdorff space $K$ and a continuous map $f: X\rightarrow K$, there is a unique continuous map $f_*: \beta(X)\rightarrow K$ such that $f_*\circ\iota = f$.
    \begin{equation*}
        \begin{tikzcd}[row sep=large, column sep=huge]
            X
            \arrow[dr, "f"']
            \arrow[r, "\iota"] &
            \beta(X)
            \arrow[d, "f_*", dashed] \\
            & K
        \end{tikzcd}
    \end{equation*}
\end{defi}

Note that a Stone-\v{C}ech compactification of a topological space $X$ is unique up to homeomorphism (but may not be unique up to equivalence), provided that it exists.
While the Stone-\v{C}ech compactification could be considered for topological spaces which are not completely regular, in this note, we only consider the Stone-\v{C}ech compactification of completely regular spaces.
\begin{thm}
    Let $X$ be a completely regular space.
    Then a Stone-\v{C}ech compactification of $X$ exists.
    Furthermore, a Stone-\v{C}ech compactification of $X$ is unique up to equivalence.
\end{thm}

The following proposition states that a compactification $Y$ of $X$ in \cref{SC-cptf_ver1}, together with the inclusion of $X$ into $Y$, is a compactification of $X$ to which every continuous map from $X$ to any compact Hausdorff space extends continuously and uniquely.
\begin{prop}[Existence part]
    Let $X$ be a completely regular space and $Y$ be a compactification of $X$ satisfying the property in \cref{SC-cptf_ver1}.
    Then, for any compact Hausdorff space $K$ and a continuous map $f: X\rightarrow K$, there is a unique continuous map $f_*: Y\rightarrow K$ extending $f$.
\end{prop}
\begin{proof}
    Remark that a compact Hausdorff space is completely regular, so $K$ embeds into $[0, 1]^I$ for some $I$.
    Let $e: K\hookrightarrow[0, 1]^I$ be such an embedding.
    To apply the result of \cref{SC-cptf_ver1}, we may consider the composition $g:=e\circ f: X\rightarrow e(K)\subset[0, 1]^I$ rather than $f$, because $g$ is into a Euclidean space so every projection of $g$ is into $\bb{R}$.
    
    For each $\alpha\in I$, let $(g_\alpha)_*: Y\rightarrow[0, 1]$ be the unique continuous extension of $g_\alpha: X\rightarrow[0, 1]$.
    Then, the map $g_*: Y\rightarrow[0, 1]^I$ defined by $g_*=((g_\alpha)_*)_{\alpha\in I}$ is the unique continuous extension of $g: X\rightarrow[0, 1]^I$.
    
    To assert that $f_*:=e^{-1}\circ g_*$ is a desired extension of $f$, it suffices to show that $f_*$ is well-defined, by justifying that $g_*(Y)$ is contained in $e(K)$.
    Because the closure of $X$ in $Y$ is $Y$ and $e(K)$ is compact,
    \begin{align*}
        g_*(Y)\subset\overline{g_*(X)}=\overline{g(X)}\subset\overline{e(K)}=e(K).
    \end{align*}
    Therefore, $f_*$ is well-defined and $f_*$ is a desired extension of $f: X\rightarrow K$ to $Y$.
    The uniqueness follows from the assumption that $K$ is a Hausdorff space.
\end{proof}

\begin{prop}[Uniqueness part]
    Let $X$ be a completely regular space.
    \begin{enumerate}
        \item[(a)]
        {
            A Stone-\v{C}ech compactification of $X$ is unique up to homeomorphism.
        }
        \item[(b)]
        {
            A compactification of $X$ which is a Stone-\v{C}ech compactification of $X$ is unique up to equivalence.
        }
    \end{enumerate}
\end{prop}
\begin{proof}
    (a) is clear, since the Stone-\v{C}ech compactification of a topological space is defined by a universal property.
    (b) can be similarly proved, if one applies the assumption that the given maps are inclusions.
\end{proof}

We end this section with defining a functor from $\textbf{Top}$ to $\textbf{CHaus}$, where $\textbf{CHaus}$ is the category of the compact Hausdorff spaces, with morphisms being the continuous maps of compact Hausdorff spaces into compact Hausdorff spaces.
For a topological space $X$, let $\beta(X)$ be the Stone-\v{C}ech compactification of $X$.
For a continuous map $f: X\rightarrow Y$, define $\beta(f): \beta(X)\rightarrow\beta(Y)$ as follows.
Let $(\beta(X), \imath)$ and $(\beta(Y), \jmath)$ be the Stone-\v{C}ech compactifications of $X$ and $Y$, respectively.
Then there is a unique continuous map $(\jmath\circ f)_*: \beta(X)\rightarrow\beta(Y)$ satisfying $(\jmath\circ f)_*\circ\imath=\jmath\circ f$; define $\beta(f)=(\jmath\circ f)_*$.
\begin{equation*}
\begin{tikzcd}[row sep=large, column sep=huge]
    X
    \arrow[r, "f"]
    \arrow[d, "\imath"']
    \arrow[dr, "\jmath\circ f"{sloped}]
    &
    Y
    \arrow[d, "\jmath"]
    \\
    \beta(X)
    \arrow[r, "\beta(f):=(\jmath\circ f)_*"']
    &
    \beta(Y)
\end{tikzcd}
\end{equation*}
Note that the above correspondence is well-defined (when choosing up to homeomorphism), and the universal property of the Stone-\v{C}ech compactification naturally explains the functoriality of $\beta$.
Hence, one may argue as follows.
\begin{prop}
    The map $\beta: \textbf{Top}\rightarrow\textbf{CHaus}$ defined above is a covariant functor.
\end{prop}