\section{Compact spaces}

\begin{defi}[Compact space]
    A topological space $X$ is said to be compact if every covering of $X$ by sets open in $X$ has a finite subcover.
    If every such covering has a countable subcover, then $X$ is called a Lindel\"{o}f space.
\end{defi}
\begin{rmk}
    In fact, we can impose an alternative definition of compactness as follows:
    \begin{quotation}
        The topological space $X$ is said to be compact, if for every collection $\mc{C}$ of closed sets in $X$ with the finite intersection property, the intersection of the members of $\mc{C}$ is nonempty. 
    \end{quotation}
    The equivalence of compactness and the above property can be checked by considering an appropriate contraposition.
\end{rmk}

As connectedness, compactness of subspaces can also be argued in larger spaces.
\begin{thm}[Compactness of subspaces]
    Let $A$ be a subspace of $X$.
    Then $A$ is compact if and only if every covering of $A$ by sets open in $X$ has a finite subcover.
    (This statement is valid even if the words `compact' are replaced by `Lindel\"{o}f.')
\end{thm}
\begin{proof}
    Suppose $A$ is compact, and $\mathcal{A}=\{A_\alpha\}_\alpha$ is a covering of $A$ by sets open in $X$.
    Then the naturally induced collection $\{A_\alpha\cap A\}$ is a covering of $A$ by sets open in $A$, and we can find a finite subcover.
    The corresponding finite subcover of $\mathcal{A}$ covers $A$.

    Suppose conversely that every covering of $A$ by sets open in $X$ has a finite subcover, and let $\mathcal{A}=\{A_\alpha\}_\alpha$ be a covering of $A$ by sets in $A$.
    Since each $A_\alpha$ can be written as $A\cap O_\alpha$ for some subset $O_\alpha$ open in $X$, the collection $\{O_\alpha\}_\alpha$ has a finite subcover.
    The corresponding finite subcover of $\mathcal{A}$ covers $A$.
\end{proof}

\begin{thm}
    Regarding closedness and compactness of subspaces, the following statements holds:
    \begin{enumerate}
        \item[(a)]
        {
            Every closed subspace of a compact space is compact.
        }
        \item[(b)]
        {
            Every Hausdorff space is compactly normal.
            In other words, any two disjoint compact subspaces can be separated by (disjoint) neighborhoods.
        }
        \item[(c)]
        {
            Every compact subspace of a Hausdorff space is closed.
            Hence, in a compact Hausdorff space, closedness and compactness of subspaces coincide.
            (In particular, every compact Hausdorff space is normal.)
        }
    \end{enumerate}
\end{thm}
\begin{proof}
    \hangindent=0.65cm
    \noindent(a)
    Let $X$ be a compact space and $Y$ be a closed subspace of $X$.
    Let $\mc{A}=\{A_\alpha\}_\alpha$ be a covering of $Y$ by sets open in $X$.
    Because the collection $\mc{A}\cup\{X\setminus Y\}$ is also an open covering of $X$, the collection has a finite subcover.
    Among them, excluding $X\setminus Y$ gives a finite subcover of $\mc{A}$ which covers $Y$.

    \noindent(b)
    We first prove that a Hausdorff space $X$ is compactly regular.
    Let $p$ be a point of $X$ and $A$ be a compact subspace of $X$ not containing $p$.
    For each $a\in A$, let $U_a$ and $V_a$ be disjoint neighborhoods of $a$ and $p$, respectively.
    Since $A$ can be covered by finitely many $U_a$'s, the union of such $U_a$'s and the intersection of the corresponding $V_a$'s are disjoint neighborhoods of $A$ and $p$, respectively.
    Thus, $X$ is compactly regular.

    To show that $X$ is compactly normal, let $A$ and $B$ be disjoint compact subspaces of $X$.
    For each $a\in A$, let $U_a$ and $V_a$ be disjoint neighborhoods of $a$ and $B$, respectively.
    Since $A$ can be covered by finitely many $U_a$'s, the union of such $U_a$'s and the intersection of the corresponding $V_a$'s are disjoint neighborhoods of $A$ and $B$, respectively.
    Therefore, $X$ is compactly normal.
    
    \noindent(c)
    If $A$ is a compact subspace of a Hausdorff space $X$, then (b) implies that $X\setminus A$ is open.
    Thus, a compact subspace of a Hausdorff space is closed.
    The desired coincidence follows from (a), and the desired normality follows readily.
\end{proof}

\begin{prop}
    A continuous image of a compact space is compact.
\end{prop}
\begin{proof}
    \color{brown}Left as an exercise.\color{black}
\end{proof}

\begin{prop}
    Let $f: X\rightarrow Y$ be a bijective continuous map.
    If $X$ is compact and $Y$ is Hausdorff, then $f$ is a homeomorphism.
\end{prop}
\begin{proof}
    If $A$ is a closed subspace of $X$, then $A$ is compact, so $f(A)$ is compact (and closed in $Y$).
\end{proof}

\begin{thm}\label{products of compact spaces}
    The product of finitely many compact spaces is compact.\footnote{In fact, this theorem extends to arbitrary products of compact spaces, which is called the Tychonoff's theorem.
    The proof of the Tychonoff's theorem will not be introduced in this note.}
\end{thm}
When proving the above theorem, the first version of the tube lemma will be used.
\begin{lem}\label{tube lemmas}
    \begin{enumerate}
        \item[(a)]
        {
            (The tube lemma)
            Consider the product space $X\times Y$, where $Y$ is a compact space.
            If $N$ is an open subspace of $X\times Y$ containing a slice $x_0\times Y$ for some $\{x_0\}\in X$, then $N$ contains a tube $W\times Y$ about $x_0\times Y$, where $W$ is a neighborhood of $x_0$ in $X$.
        }
        \item[(b)]
        {
            (A generalized tube lemma)
            Let $A$ and $B$ be compact subspaces of $X$ and $Y$, respectively, and let $N$ be an open subspace of $X\times Y$ containing $A\times B$.
            Then, there are neighborhoods $U$ of $A$ in $X$ and $V$ of $B$ in $Y$, respectively, such that $A\times B\subset U\times V\subset N$.
            ((a) readily follows from (b) when $A=\{x_0\}$ and $B=Y$.)
        }
    \end{enumerate}
\end{lem}
\begin{proof}[Proof of \cref{tube lemmas}]
    \hangindent=0.65cm
    (a)
    Since $N$ is open, for each point $(x_0, y)\in\{x_0\}\times Y$, there is a neighborhood $A_y\times B_y$ of the point $(x_0, y)$ contained in $N$, where $A_y$ and $B_y$ are open in $X$ and $Y$, respectively.
    Because $\{x_0\}\times Y$ is homeomorphic to $Y$, $\{x_0\}\times Y$ is compact, so the slice can be covered by finitely many above basis members.
    For convinience, write such members as $A_i\times B_i$ for $i=1, \cdots, n$.
    Letting $W=\bigcap_{i=1}^n A_i$, we have $\{x_0\}\times Y\subset W\times Y\subset N$.
    
    \noindent(b)
    Since $N$ is open, for each point $(a, t)\in\{a\}\times B$, there is a neighborhood $A_t^a\times B_t^a$ of the point $(a, t)$ contained in $N$, where $A_t^a$ and $B_t^a$ are open in $X$ and $Y$, respectively.
    Because the slice $\{a\}\times B$ is compact for each $a\in A$, the slice can be covered by finitely many neighborhoods; denote each of such neighborhoods by $A_i^a\times B_i^a$ for $i=1, \cdots, n(a)$.
    Finally, define
    \begin{align*}
        U^a:=\bigcap_{i=1}^{n(a)}{A_i^a},\quad V^a:=\bigcup_{i=1}^{n(a)}{B_i^a}.
    \end{align*}
    Then, $U^a\times V^a$ is an open subset of $X\times Y$ such that $\{a\}\times Y\subset U^a\times V^a\subset N$.
    Since $A$ is compact, there is a finite subcollection of $\{U^a\}_{a\in A}$ covering $A$; write it as $\{U^1, \cdots, U^k\}$.
    Finally, define
    \begin{align*}
        U:=\bigcup_{i=1}^k U^k,\quad V:=\bigcap_{i=1}^k V^k.
    \end{align*}
    Then, $U$ and $V$ are open in $X$ and $Y$, respectively, and $A\times B\subset U\times V\subset N$.
\end{proof}
\begin{proof}[Proof of \cref{products of compact spaces}]
    Let $X, Y$ be compact spaces, and $\mc{A}$ be an open cover of $X\times Y$.
    Given $x_0\in X$, because the slice $\{x_0\}\times Y$ is compact, there are finitely many members of $\mc{A}$ covering the slice $\{x_0\}\times Y$, and such members cover a tube $U(x_0)\times Y$ about $\{x_0\}\times Y$. (Here, $U(x_0)$ is assumed to be a neighborhood of $x_0$ in $X$.)
    Due to the compactness of $X$, there are finitely many points $x_1, \cdots, x_n$ such that $\bigcup_{i=1}^n U(x_i)=X$, and this justifies the compactness of $X\times Y$.
    The general result is deduced inductively.
\end{proof}

Before introducing some problems, we observe properties of ordered sets in the order topologies, regarding compactness.
\begin{prop}
    \begin{enumerate}
        \item[(a)]
        {
            Suppose $X$ is an ordered set in the order topology, and assume that $X$ is compact.
            Then $X$ has both the greatest and the least element.
        }
        \item[(b)]
        {
            (Extreme value theorem)
            Let $f: K\rightarrow Y$ be a continuous map, where $K$ is a compact space and $Y$ is an ordered set in the order topology.
            Then $f$ attains the maximum and the minimum.
        }
    \end{enumerate}
\end{prop}
\begin{proof}
    \hangindent=0.65cm
    (a)
    If $X$ has no least element, the open covering $\{(a, \infty): a\in X\}$ of $X$ contains no finite subcover of $X$, a contradiction.
    The same argument holds for the case where $X$ has no greatest element, provided that $(a, \infty)$ is replaced by $(-\infty, a)$ for each $a\in X$.

    \noindent(b)
    The subspace $f(K)$ of $Y$ is compact, and (a) implies that $f(K)$ has both the greatest and the least element.
\end{proof}
\begin{prop}
    Let $X$ be an ordered set in the order topology.
    If $X$ satisfies the least upper bound property if and only if every closed interval in $X$ is compact.
\end{prop}
\begin{proof}
    \color{red}The proof of only if part is not given in this note. \color{black}
    Assume that every closed intervel in $X$ is compact, i.e., $[a, b]$ is compact whenever $a, b\in X$ and $a<b$.
    Let $S$ be a nonempty subset of $X$ which is bounded above.
    We will show that $S$ has the supremum (the least upper bound) in $X$.
    Let $u$ be any upper bound of $S$ in $X$, and consider the set
    \begin{align*}
        U:=\bigcap_{a\in S}[a, u].
    \end{align*}
    Then $U$ is closed in $X$, so $U$ is compact.
    Because $U$ is a compact ordered set (in the order topology), $U$ has the least element $M$.
    
    We now show that $M$ is the supremum of $S$.
    It is obvious that $M$ is an upper bound of $S$ in $X$; otherwise, there is an element $b\in S$ with $M<b$, which implies $M\notin U$, a contradiction.
    Assume that $M$ is not the supremum of $S$.
    Then there is an upper bound $v$ of $S$ such that $v<M$.
    Because $v\geq a$ for all $a\in S$, we have $v\in U$, so $M\leq v$, a contradiction.
\end{proof}

The first problem deals with the distance between a point and a subspace and neighborhoods of subspaces in a metric space.
In particular, (c) implies that the original definition of the $\epsilon$-neighborhood of a subspace and our intuition coincide.
\begin{prob}
    Let $(X, d)$ be a metric space and $A$ be a nonempty subset of $A$.
    \begin{enumerate}
        \item[(a)]
        {
            Show that $\dist{d}{x}{A}=0$ if and only if $x\in\overline{A}$.
        }
        \item[(b)]
        {
            Show that if $A$ is compact, then $\dist{d}{x}{A}=d(x, a)$ for some $a\in A$, whenever $x\in X$.
        }
        \item[(c)]
        {
            Define the $\epsilon$-neighborhood of $A$ in $X$ to be the set
                \begin{align*}
                    U(A, \epsilon):=\{x\in X:\dist{d}{x}{A}<\epsilon\}.
                \end{align*}
            Show that $U(A, \epsilon)$ equals the union of the open balls $B_d(a, \epsilon)$ for $a\in A$.
        }
        \item[(d)]
        {
            Assume that $A$ is compact, and let $U$ be an open set in $X$ containing $A$.
            Show that $U$ contains an $\epsilon$-neighborhood of $A$ for some $\epsilon>0$.
        }
        \item[(e)]
        {
            Show that the result in (d) need not hold if $A$ is closed but not compact.
        }
    \end{enumerate}
\end{prob}
\begin{sol}
    \begin{enumerate}
        \item[(a)]
        {    
            Almost clear.
            If $x\in\overline{A}$, then whenever $k>0$, there is a point $a\in A$ such that $d(x, a)<k$.
            Hence, $\dist{d}{x}{A}=0$.
            Assuming conversely, whenever $k>0$, there is a point $a\in A$ such that $d(x, a)<k$, so $B_d(x, k)$ intersects $A$.
            Hence, $x\in\overline{A}$.
        }
        \item[(b)]
        {
            Given a point $x\in X$, define a function $f_x: A\rightarrow[0, \infty)$ by $f_x(a)=d(x, a)$ for $a\in A$.
            Since $|f_x(a)-f_x(b)|\leq d(a, b)$ whenever $a, b\in A$, $f_x$ is (uniformly) continuous.
            Because the domain $A$ of $f_x$ is a compact space, $f_x$ attains the minimum.
            Furthermore, $\dist{d}{x}{A}=\inf\{d(x, a): a\in A\}=\min\{f_x(a): a\in A\}$, so $\dist{d}{x}{A}=d(x, a)$ for some $a\in A$.
        }
        \item[(c)]
        {
            It is clear that every $\epsilon$-ball with the center in $A$ is contained in the $\epsilon$-neighborhood of $A$, so one inclusion is obvious.
            Suppose $u\in U(A, \epsilon)$ so that $\dist{d}{u}{A}<\epsilon$.
            It implies that the $\epsilon$-ball centered at $u$ intersects $A$ at a point $a\in A$, so the $\epsilon$-ball centered at $a$ contains $u$.
            Therefore, $U(A, \epsilon)$ is contained in the union of the $\epsilon$-balls with centers in $A$.
        }
        \item[(d)]
        {    
            Let $f: A\rightarrow \bb{R}$ be a function defined by $f(a)=\dist{d}{a}{X\setminus U}$.
            \begin{itemize}
                \item
                {
                    $f$ is continuous; for $a, b\in A$, we have $f(a)\leq d(a, x)\leq d(a, b)+d(b, x)$ for all $x\in X\setminus U$, so $f(a)\leq d(a, b)+f(b)$, from which, by symmetry, we obtain $|f(a)-f(b)|\leq d(a, b)$.
                }
                \item
                {
                    $f(a)>0$ for all $a\in A$; otherwise, if $f(p)=0$ for some $p\in A$, then $p\in\ol{X\setminus U}=X\setminus U$, a contradiction.
                }
            \end{itemize}
            Because $A$ is compact, $f$ attains the minimum $\delta>0$.
            Hence, for example, if $\epsilon=\delta/2$, then $U(A, \epsilon)$ is contained in $U$.
        }
        \item[(e)]
        {
            The $x$-axis in the $xy$-plane is closed but not compact, and $R=\{(x, y)\in\bb{R}^2: y<e^{-|x|}\}$ is open and contains the $x$-axis.
            However, no neighborhood of the $x$-axis is contained in $R$.
        }
    \end{enumerate}
\end{sol}

\begin{prob}
    Let $X$ be a compact Hausdorff space.
    Let $\mc{A}$ be a collection of closed connected subsets of $X$ that is simply ordered by proper inclusion.
    Then $Y:=\bigcap_{A\in\mc{A}} A$ is connected.
\end{prob}
\begin{sol}
    Suppose $Y$ is not connected, and let $(C, D)$ be a separation of the intersection $Y=\bigcap_{A\in\mc{A}}A$.
    Because $C$ and $D$ are closed in $Y$ and $Y$ is closed in $X$, $C$ and $D$ are closed in $X$.
    Since $X$ is normal, there are disjoint neighborhoods $U$ and $V$ of $C$ and $D$, respectively.
    
    Assume that $A\setminus (U\sqcup V)$ is empty for some $A\in\mc{A}$.
    Then $A\subset U\sqcup V$ and both $A\cap U, A\cap V$ are nonempty, for $Y\subset A$.
    Thus, $A$ has a separation $(A\cap U, A\cap V)$, a contradiction; $A\setminus(U\sqcup V)$ is nonempty (i.e., $A\not\subset U\sqcup V$) for all $A\in\mc{A}$.
    
    The collection $\mc{C}:=\{A\setminus(U\sqcup V): A\in\mc{A}\}$ is a simply ordered collection of sets closed in $X$, and $\mc{C}$ satisfies the finite intersection property.
    Because $X$ is compact, the intersection of members in $\mc{C}$ is nonempty, implying that $Y\not\subset U\sqcup V$, a contradiction.
    Therefore, $Y$ is connected.
\end{sol}