\section{Countability and separation axioms}

\subsection{Definitions regarding countability and separation axioms}
\begin{defi}[Countability axioms]
    Let $X$ be a topological space.
    \begin{enumerate}
        \item[(a)] $X$ is said to have a countable base at the point $x\in X$, if there is a countable collection $\{B_n\}_{n\in\bb{N}}$ of open subsets of $X$ with the following property:
            \begin{center}
                For each open set $A\subset X$ there is a positive integer $n$ such that $B_n$ is contained in $A$.
            \end{center}
        \item[(b)] $X$ is said to be a first-countable space (or to be first countable) if every point of $X$ has a countable base.
        \item[(c)] $X$ is said to be a second-countable space (or to be second countable) if $X$ has a countable basis.
    \end{enumerate}
    Remark that the second countability implies the first countability.
\end{defi}
\begin{rmk}[A stronger property of second-countability]
    The definition of the second-countability states that a space has a countable basis.
    In fact, if the space $X$ is second-countable and $\mc{C}$ is a basis of $X$, then $\mc{C}$ has a countable subcollection which is a basis of $X$.
\end{rmk}
\begin{proof}
    Let $\mc{B}=\{B_n\}_{n\in\bb{N}}$ be a coundtable basis of $X$.
    For each $m, n\in\bb{N}$, whenever it is possible, choose a member $C_{m, n}\in\mc{C}$ such that $B_m\subset C_{m, n}\subset B_n$.
    Indeed, for each $n\in\bb{N}$, there is such index $m\in\bb{N}$; for each point $p\in X$ and an index $n\in\bb{N}$ such that $p\in B_n$, there is a basis member $\gamma_n\in\mc{C}$ such that $p\in\gamma_n\subset B_n$, and there is an index $m\in\bb{N}$ such that $p\in B_m\subset\gamma_n$.
    
    We first show that the collection $\mc{C}^*:=\{C_{m, n}\}_{m, n}$ is a countable basis of $X$.
    The countability of $\mc{C}^*$ is obvious, and the observation in the preceeding paragraph implies that $\mc{C}^*$ covers $X$.
    Assume that a point $p\in X$ belongs to two members $C_{m, n}, C_{j, k}\in\mc{C}^*$.
    Let $C_0$ be a basis member of $\mc{C}$ such that $p\in C_0\in C_{m, n}\cap C_{j, k}$, and let $l\in\bb{N}$ be an index such that $p\in B_l\subset C_0$.
    Using the idea introduced in the preceeding paragraph again, we can find an integer $i\in\bb{N}$ such that $p\in B_i$ and $C_{i, l}$ exists.
    The member $C_{i, l}$ is one of a basis member in $\mc{C}^*$ we have been seek to find.

    Cleraly, since $\mc{C}^*$ is contained in $\mc{C}$, the topology generated by $\mc{C}^*$ is coarser than the topology on $X$.
    Conversrly, given $x\in X$ with a basis member $B_n\in\mc{B}$ containing $x$, there is a basis member $C_{l, n}\in\mc{C}$ containing $x$ for some $l\in\bb{N}$.
    Therefore, $\mc{C}^*$ is a countable basis of $X$ which is contained in $\mc{C}$.
    \end{proof}
\begin{defi}[Separation axioms]
    Let $X$ be a topological space.
    $X$ is said to be a Kolomogrov space (or a $T_1$ space) if every finite subset of $X$ is closed in $X$.
    In the rest definitions, assume $X$ is a Kolomogrov space.
    \begin{enumerate}
        \item[(a)]
        {
            $X$ is said to be a Hausdorff space (or a $T_2$ space) if given two distinct points $a$ and $b$ in $X$, there are disjoint neighborhoods of $a$ and $b$.
        }
        \item[(b)]
        {
            $X$ is said to be a regular space (or a $T_3$ space) if given a point $a\in X$ and a nonempty closed subset $B\subset X$ not containing $a$, there are disjoint neighborhoods of $a$ and $B$.
        }
        \item[(c)]
        {
            $X$ is said to be a completely regular space (or a $T_{3\frac{1}{2}}$ space) if given a point $a\in X$ and a nonempty closed subset $B\subset X$ not containing $a$, there is a continuous function $f: X\rightarrow[0, 1]$ such that $f(a)=1$ and $f(B)=\{0\}$.
        }
        \item[(d)]
        {
            $X$ is said to be a normal space (or a $T_4$ space) if given two nonempty disjoint subsets in $X$ there are disjoint neighborhoods of those closed subsets.
        }
    \end{enumerate}
\end{defi}
\begin{rmk}[Alternative definitions for some separabilities in terms of locality]
    Suppose $X$ is a $T_1$ space.
    Some separabilities on $X$ may be defined alternatively as following statements written in terms of locality:
    \begin{enumerate}
        \item[(a)]
        {
            $X$ is a regular space if and only if given a point $a\in X$ with its neighborhood $U$ in $X$, there is a neighborhood of $a$ in $X$ whose closure in $X$ is contained in $U$.
        }
        \item[(b)]
        {
            $X$ is a completely regular space if and only if given a point $p$ of $X$ and its neighborhood $U$ in $X$, there is a continuous function $f: X\rightarrow[0, 1]$ such that $f(p)=1$ and $f(X\setminus U)=\{0\}$.
        }
        \item[(c)]
        {
            $X$ is a normal space if and only if given a nonempty closed subset $B\subset X$ with its neighborhood $U$ in $X$, there is a neighborhood of $B$ in $X$ whose closure in $X$ is contained in $U$.
        }
    \end{enumerate}
    \color{brown}Proving the above equivalences is left as an exercise.\color{black}
\end{rmk}

\subsection{Basic properties of countabilities and separabilities}
As one can expect from the fact that a basis of the topology plays a major role in theory, second-countability is a strong condition.
\begin{prop}\label{2nd ctbl implies}
    Suppose $X$ is a second-countable space.
    \begin{enumerate}
        \item[(a)] $X$ is a Lindel\"{o}f space, i.e., every open covering of $X$ has a countable subcover.
        \item[(b)] $X$ is a separable space, i.e., $X$ has a coundtable dense subset.
    \end{enumerate}
\end{prop}
\begin{proof}
    \color{brown}Left as an exercise.\color{black}
\end{proof}
The converse implications in \cref{2nd ctbl implies} are sometimes valid.
One of such a particular case is when the space is metrizable, which is also one of a desired property.
\begin{prop}
    For metrizable spaces, second countability, separability, and being a Lindel\"{o}f space coincide.
    In other words,
    \begin{enumerate}
        \item[(a)]
        {
            every metrizable Lindel\"{o}f space is second countable.
        }
        \item[(b)]
        {
            every metrizable separable space is second countable.
        }
    \end{enumerate}
\end{prop}
\begin{proof}
    Let $X$ be a metrizable space, and let $d$ be a metric on $X$ which induces the topology on $X$.
    \begin{enumerate}
        \item[(a)]
        {
            For each $n\in\bb{N}$, the open cover $\{B_d\left(x, n^{-1}\right): x\in X\}$ has a countable subcover; let each member of a countable subcover be denoted by $B_k^n=B_d(x_k^n, n^{-1})$ with $k\in\bb{N}$.
            We want to show that the collection $\mc{B}:=\{B_k^n\}_{n, k\in\bb{N}}$ is a countable basis of the topology on $X$.
            Clearly, $\mc{B}$ is an open cover of $X$.
            If a point $p\in X$ is contained in two members $B_{k_1}^{n_1}$ and $B_{k_2}^{n_2}$, let $r$ be the positive number which is the minimum among the following four positive numbers:
            \begin{align*}
                d(p, x_{k_1}^{n_1}),\quad\dfrac{1}{n_1}-d(p, x_{k_1}^{n_1}),\quad d(p, x_{k_2}^{n_2}),\quad\dfrac{1}{n_2}-d(p, x_{k_2}^{n_2}).
            \end{align*}
            Suppose $n$ is large so that $\dfrac{2}{n}<r$, and let $j$ be an index such that $B_j^n$ contains $p$.
            Then $B_j^n$ is a member of $\mc{B}$ such that $p\in B_j^n\subset B_{k_1}^{n_1}\cap B_{k_2}^{n_2}$.
            Thus, $\mc{B}$ is a basis of a topology on $X$.
            It is clear that the topology on $X$ generated by $\mc{B}$ is coarser than the topology on $X$.
            The converse inclusion is also clear, when considering how $\mc{B}$ consists of. \color{brown}(Proving details are left as an exercise.)\color{brown}
        }
        \item[(b)]
        {
            Let $D=\{a_n\}_{n\in\bb{N}}$ be a countable dense subset of $X$.
            The following countable collection
            \begin{align*}
                \left\{B_d(a_n, 1/k): k\in\bb{N}\right\}
            \end{align*}
            is a basis of the topology on $X$. \color{brown}(Think why.)\color{black}
        }
    \end{enumerate}
    Hence, when the space is metrizable, Lindel\"{o}f condition or separability implies second-countability.
\end{proof}

We now introduce some inheritance properties.
\begin{prop}[Inheritance of countabilities]
    Some countabilities are preserved as follows:
    \begin{enumerate}
        \item[(a)] First and second countabilities are inherited to subspaces and countable product spaces.
        \item[(b)] Every closed subspace of a Lindel\"{o}f space is also a Lindel\"{o}f space.
        \item[(c)] Every open subspace of a separable space is also a separable space.
    \end{enumerate}
\end{prop}
\begin{proof}
    In this note, we only prove part (c).
    Let $D$ be a countable dense subset of the separable space $X$, and let $A$ be an open subset of $X$.
    We want to check if the closure of $A\cap D$ in $A$ is $A$; in other words, if $\overline{A\cap D}\cap A=A$.
    Suppose $a\in A$ and $U$ is a neighborhood of $a$ contained in $A$.
    Then $U$ intersects $D$, hence $U$ intersects $A\cap D$.
\end{proof}

\begin{prop}[Inheritance of separabilities]
    \begin{enumerate}
        \item[(a)]
        {
            A subspace of a Hausdorff (regular, completely regular) space is a Hausdorff (regular, completely regular) spaces.
        }
        \item[(b)]
        {
            A product of Hausdorff (regular, completely regular) spaces is a Hausdorff (regular, completely regular) spaces.
        }
        \item[(c)]
        {
            For the normal space $X$, a closed subspace of $X$ is normal.
        }
        \item[(d)]
        {
            If the product space $\prod_{\alpha\in A}X_\alpha$ is a Hausdorff (regular, normal) space, then so is each $X_\alpha$ for $\alpha\in A$.\label{converse inheritance}
        }
    \end{enumerate}
\end{prop}
\begin{proof}
    \begin{enumerate}
        \item[(a)]
        {           
            Since the result is clear when the space $X$ is a Hausdorff space, assume $X$ is a regular space and $Y$ is a subspace of $X$.
            Let $x$ be a point of $Y$ and $B$ be a closed subspace of $Y$ not containing $x$.
            Because $Y\cap\overline{B}=B$, $x\notin\overline{B}$.
            Using regularity, $x$ and $\overline{B}$ can be separated by two disjoint sets $U$ and $V$ open in $X$.
            Thus, $Y\cap U$ and $Y\cap V$ form a desired pair.
        }
        \item[(b)]
        {
            We also prove for regular case only.
            Suppose $\{X_\alpha\}_{\alpha\in A}$ is a collection of regular spaces, and let $X=\prod_{\alpha\in A}X_\alpha$.
            Suppose $p\in X$ and $U\subset X$ is a neighborhood of $p$ in $X$.
            Let $\prod_{\alpha\in A}V_\alpha\subset X$ be a basis member containing $p$ and contained in $U$.
            (Here, $\{\alpha\in A:V_\alpha\neq X_\alpha\}=\{\alpha_1, \cdots, \alpha_n\}$.)
            For each $\alpha\in A$, use regularity to find a neighborhood $W
            _\alpha$ of $p_\alpha$ in $X_\alpha$ whose closure in $X_\alpha$ is contained in $V_\alpha$, with $W_{\alpha_i}=X_{\alpha_i}$ for $i=1, \cdots, n$.
            Then, $\prod_{\alpha\in A}W_\alpha$ is a neighborhood of $p$ in $X$, and its closure satisfies
            \begin{align*}
                \overline{\prod_{\alpha\in A}W_\alpha}=\prod_{\alpha\in A}\overline{W_\alpha}\subset\prod_{\alpha\in A}V_\alpha\subset U.
            \end{align*}
            Therefore, the product space $X$ is also a regular space.
            }
        \item[(c)]
        {    
            Suppose $Y$ is a closed subspace of the normal space $X$.
            If $A$ and $B$ are disjoint closed subspaces of $Y$, then they are closed in $X$, too.
            By normality, there are disjoint open subspaces $U$ and $V$ in $X$ containing $A$ and $B$, respectively.
            So $Y\cap U$ and $Y\cap V$, which are open subspaces of $Y$, separate $A$ and $B$.
            Therefore, a closed subspace of a normal space is normal.
        }
    \end{enumerate}
    The proof for complete regular spaces and the proof of (d) are not given in this section.
\end{proof}

\subsection{Examples of regular spaces}
\begin{exmp}[Ordered spaces are regular]
    Let $X$ be an ordered space, and let $U=(a, b)$ be a basis member in $X$ containing $p\in X$.
    \begin{enumerate}
        \item[(\romannumeral 1)]
        {
            Suppose $U$ consists of only one point $p$.
            Then, $U$ is both open and closed in $X$, so we may choose $V=U$.
        }
        \item[(\romannumeral 2)]
        {
            Suppose that $(a, p)$ or $(p, b)$ is empty.
            Without loss of generality, we assume that $(p, b)$ is empty and $(a, p)$ is nonempty.
            Choose a point $a'\in(a, p)$ and let $V=(a', b)=(a', p]$.
            Any point $x\in X$ with $x>p$ does not belong to the closure of $V$ in $X$; $(p, \infty)$ does not intersect $U$.
            Any point $x\in X$ with $x<a'$ does not belong to the closure of $V$ in $X$; $(-\infty, a')$ does not intersect $V$.
            Hence, the closure of $V$ in $X$ is contained in $[a', p]\subset U$.
        }
        \item[(\romannumeral 3)]
        {
            Suppose $(a, p)$ and $(p, b)$ are nonempty.
            Let $a'$ and $b'$ be points of $X$ in $(a, p)$ and $(p, b)$, respectively, and let $V=(a', b')$.
            Clearly, $V$ is a neighborhood of $p$ in $X$, and it is now easy to check that the closure of $V$ in $X$ is contained in $U$.
        }
    \end{enumerate}
    By from (\romannumeral 1) to (\romannumeral 3), every ordered space is a regular space.
    In fact, it is known that every ordered space is a normal space, whose proof will not be introduced in this note.
\end{exmp}

\begin{exmp}[Locally compact Hausdorff spaces are regular]
    Suppose $X$ is a locally compact Hausdorff space.
    Without loss of generality, we may assume $X$ is not compact. (You will see why.)
    Let $Y$ be the one-point compactification of $X$.
    Since $Y$ is a compact Hausdorff space, $Y$ is normal and regular, as illustrated in the following subsection.
    Hence, $X$, a subspace of $Y$, is also regular.
\end{exmp}

\subsection{Examples of normal spaces}
\begin{thm}
    Lindel\"{o}f regular spaces are normal.
\end{thm}
\begin{proof}
    Let $A$ and $B$ be disjoint closed subspaces of $X$.
    For each $a\in A$, let $S_a$ be a neighborhood of $a$ in $X$ contained in $X\setminus B$.
    Using regularity, let $U_a$ be a neighborhood of $a$ in $X$ whose closure in $X$ is contained in $S_a$.
    And construct $V_b$ for each $b\in B$ as we constructed $U_a$ for each $a\in A$.

    Even if $\{U_a\}_{a\in A}$ and $\{V_b\}_{b\in B}$ are open covers of $A$ and $B$, respectively, the unions of the members in each collection need not be disjoint.
    (By drawing a picture on a sketchbook) we may set
    \begin{align*}
        F_n:=U_n\setminus\bigcup_{k=1}^n\overline{V_k},\quad G_n:=V_n\setminus\bigcup_{k=1}^n\overline{U_k}
    \end{align*}
    for each $n\in\bb{N}$, and
    \begin{align*}
        F:=\bigcup_{n=1}^\infty F_n,\quad G:=\bigcup_{n=1}^\infty G_n.
    \end{align*}
    We wish $F$ and $G$ to be disjoint neighborhoods of $A$ and $B$, respectively.
    Clearly, $F$ and $G$ are neighborhoods of $A$ and $B$, since
    \begin{align*}
        F=\bigcup_{n=1}^\infty U_n\setminus\bigcup_{n=1}^\infty\overline{V_n}\supset A,\quad G=\bigcup_{n=1}^\infty V_n\setminus\bigcup_{n=1}^\infty\overline{U_n}\supset B.
    \end{align*}
    Also, $F\cap G=\varnothing$; otherwise, $F_n\cap G_k\neq\varnothing$ for some $n, l\in\bb{N}$.
    Without loss of generality, we may assume $n\leq k$.
    If $p\in F_n\cap G_k$, then $p\in F_n\subset U_n$; on the other hand, $G_k$ does not intersect $\overline{U_n}$, hence $p\notin G_k$.
    
    Therefore, $F$ and $G$ separate $A$ and $B$ and $X$ is a normal space.
\end{proof}

\begin{thm}
    Metrizable spaces are normal.
\end{thm}
\begin{proof}
    Let $X$ be a metrizable space and $d$ be a metric on $X$ inducing the topology on $X$.
    Let $A$ and $B$ be nonempty and disjoint closed subspaces of $X$.
    For each $a\in A$ and $b\in B$, let $s_a$ and $t_b$ be positive real numbers such that $B_d(a, s_a)\subset X\setminus B$ and $B_d(b, t_b)\subset X\setminus A$.
    Define
    \begin{align*}
        U:=\bigcup_{a\in A}B_d(a, s_a/3),\quad V:=\bigcup_{b\in B}B_d(b, t_b/3).
    \end{align*}
    Then $U$ and $V$ are neighborhoods of $A$ and $B$ in $X$.
    If $U\cap V$ is nonempty, then $B_d(a, s_a/3)\cap B_d(b, t_b/3)$ contains a point $z\in X$.
    We then have the following inequality:
    \begin{align*}
        d(a, b)\leq d(a, z)+d(z, b)<\dfrac{s_a}{3}+\dfrac{t_b}{3}<\max\left\{s_a, t_b\right\},
    \end{align*}
    so $b\in B_d(a, s_a)$ or $a\in B_d(b, t_b)$, a contradiction.
    Therefore, $U\cap V$ is empty, proving that $X$ is normal.
\end{proof}

\begin{thm}
    Well-ordered sets are normal. (In fact, every ordered space is normal.)
\end{thm}
\begin{proof}
    Let $X$ be a well-orderd set.

    \noindent(Step 1: Proving that every subspace of $X$ of the form $(x, y]$ is open in $X$)\newline\indent
    Let $A=(x, y]$ be a subset of $X$.
    Then $X\setminus A=(-\infty, x]\sqcup(y, \infty)$; because $(y, \infty)$ has the least element $y'\in X$, $X\setminus A=(-\infty, x]\sqcup[y', \infty)$ is closed in $X$.

    \noindent(Step 2: Proving that $X$ is normal)\newline\indent
    Let $A$ and $B$ be nonempty disjoint closed subspace of $X$.
    \begin{enumerate}
        \item[(\romannumeral 1)]
        {
        Assume that neither of them contains the least element $m$ of $X$.
        For each $a\in A$, there is a neighborhood $(x_a, a]$ of $a$ in $X$ not intersecting $B$; for each $b\in B$, there is a neighborhood $(y_b, b]$ of $b$ in $X$ not intersecting $A$.
        Define
        \begin{align*}
            U:=\bigcup_{a\in A}(x_a, a],\quad V:=\bigcup_{b\in B}(y_b, b].
        \end{align*}
        They are open in $X$ and $U$ covers $A$, and $V$ covers $B$.
        If $U$ and $V$ are not disjoint, $(x_a, a]\cap(y_b, b]$ is nonempty for some $a\in A$ and $b\in B$,
        Without loss of generality, we may assume $a<b$; then we have $a\in(y_b, b]$, a contradicton.
        }
        \item[(\romannumeral 2)]
        {
            Suppose, without loss of generality, $A$ contains $m$.
            Because the singletone $\{m\}=[m, m]=(-\infty, m]$ is not only closed but also open in $X$, $A\setminus\{m\}$ is also closed in $X$.
            By the preceeding part, there are disjoint neighborhood $U$ and $V$ of $A\setminus\{m\}$ and $B$, respectively.
            Then, $U\cup\{m\}$ and $V\setminus\{m\}$ are disjoint neighborhoods of $A$ and $B$, respectively.
        }
    \end{enumerate}
    Therefore, every well-ordered space is normal.
\end{proof}

\begin{exmp}[$\bb{R}_l$ is normal]
    Suppose $A$ and $B$ are disjoint closed subspaces of $\bb{R}_l$.
    For each $a\in A$, let $[a, x_a)$ be a neighborhood of $a$ in $\bb{R}_l$ contained in $\bb{R}_l\setminus B$; for each $b\in B$, let $[b, y_b)$ be a neighborhood of $b$ in $\bb{R}_l$ contained in $\bb{R}_l\setminus A$.
    Now set
    \begin{align*}
        U:=\bigcup_{a\in A}[a, x_a),\quad V:=\bigcup_{b\in B}[b, y_b).
    \end{align*}
    To show normality, it remains to show that $U$ and $V$ are disjoint.
    If $U\cap V\neq\varnothing$, then $[a, x_a)\cap[b, y_b)$ is nonempty for some $a\in A$ and $b\in B$.
    Since $a\neq b$, we may assume $a<b$; then $a<b<x_a<y_b$, and this implies that the point $b$ of $B$ is contained in $[a, x_a)$, a contradiction.
    Therefore, $\bb{R}_l$ is a normal space.
\end{exmp}

\begin{exmp}[Compact Hausdorff spaces are normal]
    Because closedness and compactness coincide in compact Hausdorff spaces and compact Hausdorff spaces are compactly normal, compact Hausdorff spaces are normal.
\end{exmp}