\section{Cyclic extensions}

Starting from this section, we study the splitting field for $t^n-a\in F[t]$ over $F$ under some conditions, due to some technical problems.
\begin{obs}
    Let $F$ be a field.
    Then $t^n-a\in F[t]$ is separable whenever $a\neq 0$ and $\ch{F}$ does not divide $n$.
\end{obs}
Note that we are mainly interested in (finite) Galois extensions and the condition that $\ch{F}$ does not divide $n$ (together with $a\in F\setminus\{0\}$) is the separability condition.
Hence, we are interested in the splitting field for $t^n-a\in F[t]$ over $F$ under the following assumption:
\begin{center}
    Assumption: $F$ is a field and $\ch{F}$ does not divide $n$.
\end{center}
In particular, when $\ch{F}=0$, then $t^n-a$ is separable whenever $n\in\bb{N}$ and $a\neq 0$.

\begin{prop}\label{radical -> cyclic}
    Let $F$ be a field and $n$ be a positive integer which is not divisible by $\ch{F}$, and assume that $F$ contains an $n$-th root of unity.
    If $a\in F$ and $\alpha$ is a root of $t^n-a\in F[t]$, then
    \begin{enumerate}
        \item[(a)]
        {
            $F(\alpha)$ is the splitting field for $t^n-a$ over $F$.
        }
        \item[(b)]
        {
            $F(\alpha)/F$ is a cyclic extension and $\gal{F(\alpha)/F}$ embeds into $Z_n$.
        }
    \end{enumerate}
\end{prop}
\begin{proof}
    The proof of (b) would be sufficient.
    Without loss of generality, we may assume $a\neq 0$, for $a=0$ forces $F(\alpha)=F$.
    Since $t^n-a$ is a separable polynomial over $F$, $F(\alpha)/F$ is a finite Galois extension, and an automorphism $\sigma\in\gal{F(\alpha)/F}$ maps $\alpha$ to $\alpha\zeta^k$, where $\zeta$ is a primitive $n$-th root of unity and $k$ is an integer.
    This induces the map $\rho: \gal{F(\alpha)/F}\rightarrow \bb{Z}/n\bb{Z}$ defined by $\rho\sigma=\ol k$, which is a group monomorphism.
\end{proof}
\begin{cor}
    Let $F$ be a field and $n$ be a positive integer which is not divisible by $\ch{F}$.
    If $a\in F$ and $f(t)=t^n-a\in F[t]$, then $G_f$ is a solvable group.
\end{cor}
\begin{proof}
    The roots of $f(t)$ are $\alpha, \alpha\zeta, \cdots, \alpha\zeta^{n-1}$, where $\alpha$ is a root of $f(t)$ and $\zeta$ is a primitive $n$-th root of unity. (Remark that the separability of $f(t)$ is ensured by the assumption on $\ch{F}$ and $n$.)
    Consider the tower $F\leq F(\zeta)\leq F(\alpha, \zeta)$, and observe that both $F(\zeta)/F$ and $F(\alpha, \zeta)/F(\zeta)$ are (finite) abelian extensions.
    It now follows that $G_f$ is a solvable group.
\end{proof}

We now prove the following converse of (b) in \cref{radical -> cyclic}.
\begin{prop}
    Suppose that $n$ is a positive integer and $F$ is a field whose characteristic does not divide $n$.
    Assume that $F$ contains a primitive $n$-th root of unity.
    If $K/F$ is a finite cyclic extension of degree $n$, then there is an element $\alpha\in K$ such that $\alpha^n\in F$ and $K=F(\alpha)$ (in short, $K=F(\sqrt[n]{a})$ for some $a\in F$).
\end{prop}
\begin{proof}
    See \cref{the proof of cyclic->radical}.
\end{proof}
\begin{rmk}
    We summarize the preceeding equivalence as follows:
    \begin{quotation}
        Suppose that $n$ is a positive integer and $F$ is a field whose characteristic doen not divide $n$, and assume further that $F$ contains a primitive $n$-th root of unity.
        \begin{enumerate}
            \item[(a)]
            {
                Let $\alpha$ be a root of $t^n-a\in F[t]$.
                Then $F(\alpha)$ is the splitting field for $t^n-a$ over $F$, and $F(\alpha)/F$ is a cyclic extension of degree dividing $n$.
            }
            \item[(b)]
            {
                Conversely, if $K/F$ is a cyclic extension of degree $n$, then there is an element $\alpha\in K$ such that $K=F(\alpha)$ and $\alpha^n\in F$.
                (Hence, $K/F$ is a radical extension.)
            }
        \end{enumerate}
    \end{quotation}
\end{rmk}